---
title: "esm206_lab5"
author: "Russell Fager"
date: "10/25/2021"
output: 
  html_document: 
    toc: yes
    theme: united
    code_folding: hide
---

```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(effsize)
```

## Part 1: Confidence intervals 

In this section we wil use the `t.test()` function to calculate confidence intervals.

### Example 1

Create a single mock sample containing a number of values, then learn to use `t.test()` to return confidence intervals, and learn about in-line code referencing so our text become more reproducible, too.

```{r}
# Create a mock sample

otter_lengths <- c(30, 45, 28, 31, 34, 46, 52, 29)

# Find the confidence interval

otter_ci<-t.test(otter_lengths)

# find standard deviation

otter_sd<-sd(otter_lengths)

```

if we took multiple samples from the population and calculated a confidence interval for each we would expect 95% of those intervals to have the population mean. 

Mean otter length is 36.9 inches $\pm$ 9.3 (mean $\pm$ sd, n=8) with a 95% confidence interval of [29.1, 44.7] inches.


#### Now, using in-line referencing

Mean otter length is `r round(otter_ci$estimate, 1)` inches with a 95% confidence interval of [`r round(otter_ci$conf.int[1], 1)`, `r round(otter_ci$conf.int[2], 1)`] inches.

## Part 2: Two sample t-tests

### Example 1

Make two mock samples and ask the question: do we have sufficient evidence to make us reject the null hypothesis that these samples were drawn from populations with the same mean?

```{r}
# make my mock samples

oak_trees <- c(20, 19, 24, 28, 34, 12, 15, 25)

pine_trees <- c(39, 26, 20, 30, 37, 22, 19, 18)

# Two-sample t-test
tree_t <- t.test(oak_trees, pine_trees)

#null hyopthesis: mean heights are the same

#i.e: trees are drawn from populations with the same mean height
```

if these are drawn from populations of the same mean height, there's a 28.7% chance of taking a random samples from each population and finding the same mean by chance. i.e: 30 out of 100 times I could get results that are at least this different.


the mean heights of oak and pine trees were (X +- Y) and (X +- Y), respectively, and the difference was not significant (t(`r tree_t$paramater`) = `r round(tree_t$statistic, 1)`, p = `r tree_t$p.value`).

there's and 'x%' chance of getting samples that are at least this different from populations with the same mean.

## Part 3: another t-test with cars data

using the `mpg` data set

### Take a look at the data 

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))

#check for remaining levels(
  unique(compact_suv$class)

  
```

Create a histogram of city mileage (cty) faceted by class starting from compact_suv.

```{r}
ggplot(data = compact_suv, aes(x = cty))+
  geom_histogram(bins = 12) +
  facet_wrap(~class)

ggplot(data = compact_suv, aes(sample = cty)) +
geom_qq() +
  facet_wrap(~ class)
```

find some summary statistics for each group

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(
    mean_city = round(mean(cty), 2),
    sd_city = round(sd(cty), 2),
    sample_size = n()
  )

car_stats
```

the t.test() function really wants vectors of things. Use the `pull()` function to pull out a vector from a data frame.

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)
```


```{r}
suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)
```

mean city gas mileages for suvs and compact cars are the same = null hypothesis 

```{r}
t.test(compact_sample, suv_sample)
```
there's a very very small chance of taking a random sample from each population and finding the same mean by chance. we reject the null hypothesis. There is a significant difference in mean city gas mileage for suv's and compact cars (t(df)) = t-statistic, p<0.001
