---
title: "esm206_lab5"
author: "Russell Fager"
date: "10/25/2021"
output: 
  html_document: 
    toc: yes
    theme: united
    code_folding: hide
---

```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(effsize)
```

## Part 1: Confidence intervals 

In this section we wil use the `t.test()` function to calculate confidence intervals.

### Example 1

Create a single mock sample containing a number of values, then learn to use `t.test()` to return confidence intervals, and learn about in-line code referencing so our text become more reproducible, too.

```{r}
# Create a mock sample

otter_lengths <- c(30, 45, 28, 31, 34, 46, 52, 29)

# Find the confidence interval

otter_ci<-t.test(otter_lengths)

# find standard deviation

otter_sd<-sd(otter_lengths)

```

if we took multiple samples from the population and calculated a confidence interval for each we would expect 95% of those intervals to have the population mean. 

Mean otter length is 36.9 inches $\pm$ 9.3 (mean $\pm$ sd, n=8) with a 95% confidence interval of [29.1, 44.7] inches.


#### Now, using in-line referencing

Mean otter length is `r round(otter_ci$estimate, 1)` inches

